---
title: "PLPA 6820 Coding Challenge 5"
author: "SA Donohoo"
date: "2025-03-20"
output:
  word_document:
    toc: true
  html_document:
    toc_float: true
  md_document:
    variant: gfm
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = FALSE, message = FALSE)
```

# Github Link
Sam Donohoo - PLPA 6820 - https://github.com/sad0046/PLPA6820_SP25

### Question 1 - Adding Data
Download two .csv files from Canvas called DiversityData.csv and Metadata.csv, and read them into R using relative file paths.
```{r Question 1, echo=TRUE}
# Load in Libraries
library(tidyverse)

# Load in Metadata Data
metaData <- read.csv("Metadata.csv", header = TRUE, na.strings = "na")

# Load in Diversity Data
divData <- read.csv("DiversityData.csv", header = TRUE, na.strings = "na")
```

### Question 2 - Combining Data
Join the two dataframes together by the common column ‘Code’. Name the resulting dataframe *alpha*.
```{r Question 2, echo=TRUE}
# Join Dataframes Together
# Name Combined Dataframe alpha
alpha <- full_join(metaData, divData, by = "Code")

head(alpha) # Print First 6 Rows
```

### Question 3 - Pielou's Evenness Index
Calculate Pielou’s evenness index: Pielou’s evenness is an ecological parameter calculated by the Shannon diversity index (column Shannon) divided by the log of the richness column. 
```{r Question 3, echo=TRUE}
# Calculate Pielou's Evenness Index = Shannon / log(richness)
alpha_even <- alpha %>%
  mutate(logRich = log(richness)) %>% # creating a new column of the Log Richness
  mutate(even = shannon/logRich) # creating a new column of Pielou's Evenness

head(alpha_even) # Print First 6 Rows
```

### Question 4 - Summarise
Using tidyverse language of functions and the pipe, use the summarise function and tell me the mean and standard error evenness grouped by crop over time.
```{r Question 4, echo=TRUE}
# Using summarise() Calculate Mean and Standard Error Grouped by Crop Over Time
alpha_average <- alpha_even %>%
  group_by(Crop, Time_Point) %>% # Grouping by Crop and Time Point
  summarise(Mean.even = mean(even), # Calculating the Mean and Standard Error for Evenness
            n = n(),
            sd.dev = sd(even)) %>%
  mutate(std.err = sd.dev/sqrt(n))

alpha_average # Print Entire Table
```
### Question 5 - Differences
Calculate the difference between the soybean column, the soil column, and the difference between the cotton column and the soil column.
```{r Question 5, echo=TRUE}
# Calculate Difference Between Soybean+Soil and Cotton+Soil
alpha_average2 <- alpha_average %>%
  select(Time_Point, Crop, Mean.even) %>% # Selecting Columns
  pivot_wider(names_from = Crop, values_from = Mean.even) %>% # Pivot to Wide Format
  mutate(diff.cotton.even = Soil - Cotton) %>% # calculates the mean per Treatment and Fungicide
  mutate(diff.soybean.even = Soil - Soybean)

alpha_average2 # Print Entire Table
```

### Question 6 - Plotting
Connecting it to plots.
```{r Question 6, echo=TRUE}
# Plot Differences using ggplot2
# Time_Point on the X, Value on the Y, Color Lines by Diff
alpha_average2_plot <- alpha_average2 %>%
  select(Time_Point, diff.cotton.even, diff.soybean.even) %>% # Selecting Columns
  pivot_longer(c(diff.cotton.even, diff.soybean.even), names_to = "diff") %>% # Pivot to Long Format
  ggplot(aes(x = Time_Point, y = value, color = diff)) + # Plot
  geom_line() +
  theme_classic() +
  xlab("Time (hrs)") +
  ylab("Difference from soil in Pielou's evenness")

alpha_average2_plot # Print Plot
```
