---
title: "Coding Challenge #5 - For Loop"
author: "SA Donohoo"
date: "2025-03-27"
output:
  md_document:
    variant: gfm
  word_document:
    toc: true
  html_document:
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = FALSE, message = FALSE)
```

# Github Link
Sam Donohoo - PLPA 6820 - https://github.com/sad0046/PLPA6820_SP25

### Question 1 - Reproducibility
Regarding reproducibility, what is the main point of writing your own functions and iterations?

**Answer**: Functions and iterations simplify the code and help to minimize copy and paste errors that occur when you manually type out or paste/edit chunks of code.

### Question 2 - Explain For Loops
In your own words, describe how to write a function and a for loop in R and how they work. Give me specifics like syntax, where to write code, and how the results are returned.

**Answer**: A function is composed of variables, code to be run using said variables, and usually a way to output the results of the function. An initial function can be written without variables in either the consule or an R Script. For Example:
```{r Question 2.1, echo=T}
# Base Function without Assigning Values to Variables
Sams.Function <- function(Variable1, Variable2) { # Inputs or Variables to Be Read. Uses {}.
  Mathing <- Variable1 + Variable2 # Code to Be Run
  return(Mathing) # Way to Output Results
}
```
Now we can assign values to the two variables and run the function.These results are then printed in the consule or can appended to a dataframe.
```{r Question 2.2, echo=T}
# Base Function without Assigning Values to Variables
Mathing <- Sams.Function(1:5,2)

# Print Results
print(Mathing)
```

**Answer**: A for loop is composed of a variable (organized in a list/array/sequence) and code to be run using said variable. Since for loops iteratively execute a block of code either a specific number of times or for each item in a list/array, an input variable can be defined before the for loop or during the for loop. Both the variable and for loop can be written in the consule or an R Script. The results of the for loop are then printed in the consule or can appended to a dataframe. For Example:
```{r Question 2.3, echo=T}
# Input Variable Organized as a Vector
Input.Variable <- 1:5

# For Loop Using Previously Assigned Variable
for (i in Input.Variable) { # 
  print(i + 5)
}

# For Loop Assigning Variable in Loop
for (i in 1:5) { # 
  print(i + 5)
}
```

### Question 3 - Adding Data
Read in the Cities.csv file from Canvas using a relative file path.
```{r Question 3, echo=TRUE}
# Load in Packages
library(ggplot2)
library(drc) 
library(tidyverse)
library(dplyr)

# Read in Data

All.Cities <- read.csv("../Raw_Data_for_Class_Exercises/Cities.csv", header = TRUE, na.strings = "NA")
```

### Question 4 - Generate a For Loop
Write a function to calculate the distance between two pairs of coordinates based on the Haversine formula (see below). The input into the function should be lat1, lon1, lat2, and lon2. The function should return the object distance_km. All the code below needs to go into the function.

#### Haversine formula
```{r eval=FALSE, echo=TRUE}
# Convert to radians
rad.lat1 <- lat1 * pi/180
rad.lon1 <- lon1 * pi/180
rad.lat2 <- lat2 * pi/180
rad.lon2 <- lon2 * pi/180

# Haversine formula
delta_lat <- rad.lat2 - rad.lat1
delta_lon <- rad.lon2 - rad.lon1
a <- sin(delta_lat / 2)^2 + cos(rad.lat1) * cos(rad.lat2) * sin(delta_lon / 2)^2
c <- 2 * asin(sqrt(a)) 

# Earth's radius in kilometers
earth_radius <- 6378137

# Calculate the distance
distance_km <- (earth_radius * c)/1000

```

#### Answer - Function
```{r Question 4, echo=TRUE}
distance <- function(lat1,lon1,lat2,lon2) {
  rad.lat1 <- lat1 * pi/180
  rad.lon1 <- lon1 * pi/180
  rad.lat2 <- lat2 * pi/180
  rad.lon2 <- lon2 * pi/180
  delta_lat <- rad.lat2 - rad.lat1
  delta_lon <- rad.lon2 - rad.lon1
  a <- sin(delta_lat / 2)^2 + cos(rad.lat1) * cos(rad.lat2) * sin(delta_lon / 2)^2
  c <- 2 * asin(sqrt(a))
  earth_radius <- 6378137
  distance_km <- (earth_radius * c)/1000

}
```

### Question 5 - A Single Distance
Using your function, compute the distance between Auburn, AL and New York City 

- Subset/filter the Cities.csv data to include only the latitude and longitude values you need and input as input to your function. 

- The output of your function should be 1367.854 km

```{r Question 5, echo=TRUE}
# Subset and Preset New York Coordinates
NewYork.lat <- All.Cities$lat[All.Cities$city == "New York"]
NewYork.lon <- All.Cities$long[All.Cities$city == "New York"]

# Subset and Preset Auburn Coordinates
Auburn.lat <- All.Cities$lat[All.Cities$city == "Auburn"]
Auburn.lon <- All.Cities$long[All.Cities$city == "Auburn"]

# Generate Empty Dataframe
Cities.Distance.NewYork <- NULL # create a null object

# Run Function
Distance.NewYork <- distance(NewYork.lat,NewYork.lon,Auburn.lat,Auburn.lon)

# Combine Empty and Output Dataframe
Cities.Distance.NewYork <- rbind.data.frame(Cities.Distance.NewYork, Distance.NewYork)

# Edit Column Name
colnames(Cities.Distance.NewYork) <- "km_distance"

# Print Values
print(Cities.Distance.NewYork)
```

### Question 6 - For Loop
Now, use your function within a for loop to calculate the distance between all other Cities in the data. Bonus point if you can have the output of each iteration append a new row to a dataframe, generating a new column of data. In other words, the loop should create a dataframe with three columns called city1, city2, and distance_km, as shown below.

```{r Question 6, echo=TRUE}
# Select All Cities That Are Not Auburn
Cities <- All.Cities[All.Cities$city != "Auburn",]

# Create a vector of City1 Names
Cities.names <- unique(Cities$city)

# Assign City2 Name
Auburn.name <- "Auburn"

# Generate Empty Dataframe
Cities.Distance.All <- NULL # create a null object

# Create for Loop that runs through all Cities and Calculates Distance to Auburn
for (i in seq_along(Cities.names)) {
  Cities.Distance.Single <- distance(Cities$lat[i],
                                   Cities$long[i],
                                   Auburn.lat,
                                   Auburn.lon) # Calculate Distance Between Two Cities
  Temp.df <- data.frame(Cities$city[i],Auburn.name,Cities.Distance.Single) # Create a Row with City1, City2, and km_Distance
  Cities.Distance.All <- rbind.data.frame(Cities.Distance.All, Temp.df) # Append to Previous Datframe
}

# Edit Column Names
colnames(Cities.Distance.All) <- c("City1", "City2", "km_distance")

# Print Values
print(Cities.Distance.All)
```

