---
title: "Coding Challenge #5 - For Loop"
author: "SA Donohoo"
date: "2025-03-27"
output:
  html_document:
    toc_float: true
  word_document:
    toc: true
  md_document:
    variant: gfm
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = FALSE, message = FALSE)
```

# Github Link
Sam Donohoo - PLPA 6820 - https://github.com/sad0046/PLPA6820_SP25

### Question 1 - Reproducibility
Regarding reproducibility, what is the main point of writing your own functions and iterations?
**Answer**: 

### Question 2 - Explain For Loops
In your own words, describe how to write a function and a for loop in R and how they work. Give me specifics like syntax, where to write code, and how the results are returned.
**Answer**

### Question 3 - Adding Data
Read in the Cities.csv file from Canvas using a relative file path.
```{r Question 3, echo=TRUE}
# Load in Packages
library(ggplot2)
library(drc) 
library(tidyverse)
library(dplyr)

# Read in Data

All.Cities <- read.csv("../Raw_Data_for_Class_Exercises/Cities.csv", header = TRUE, na.strings = "NA")
```

### Question 4 - Generate a For Loop
Write a function to calculate the distance between two pairs of coordinates based on the Haversine formula (see below). The input into the function should be lat1, lon1, lat2, and lon2. The function should return the object distance_km. All the code below needs to go into the function.

#### Haversine formula
```{r eval=F, echo=T}
# Convert to radians
rad.lat1 <- lat1 * pi/180
rad.lon1 <- lon1 * pi/180
rad.lat2 <- lat2 * pi/180
rad.lon2 <- lon2 * pi/180

# Haversine formula
delta_lat <- rad.lat2 - rad.lat1
delta_lon <- rad.lon2 - rad.lon1
a <- sin(delta_lat / 2)^2 + cos(rad.lat1) * cos(rad.lat2) * sin(delta_lon / 2)^2
c <- 2 * asin(sqrt(a)) 

# Earth's radius in kilometers
earth_radius <- 6378137

# Calculate the distance
distance_km <- (earth_radius * c)/1000

```

#### Answer - Function
```{r Question 4, echo=TRUE}
distance <- function(lat1,lon1,lat2,lon2) {
  rad.lat1 <- lat1 * pi/180
  rad.lon1 <- lon1 * pi/180
  rad.lat2 <- lat2 * pi/180
  rad.lon2 <- lon2 * pi/180
  delta_lat <- rad.lat2 - rad.lat1
  delta_lon <- rad.lon2 - rad.lon1
  a <- sin(delta_lat / 2)^2 + cos(rad.lat1) * cos(rad.lat2) * sin(delta_lon / 2)^2
  c <- 2 * asin(sqrt(a))
  earth_radius <- 6378137
  distance_km <- (earth_radius * c)/1000

}
```

### Question 5 - A Single Distance
Using your function, compute the distance between Auburn, AL and New York City 

- Subset/filter the Cities.csv data to include only the latitude and longitude values you need and input as input to your function. 

- The output of your function should be 1367.854 km

```{r Question 5, echo=TRUE}
# Subset and Preset New York
NewYork.lat <- All.Cities$lat[All.Cities$city == "New York"]
NewYork.lon <- All.Cities$long[All.Cities$city == "New York"]

# Subset and Preset Auburn Coordinates
Auburn.lat <- All.Cities$lat[All.Cities$city == "Auburn"]
Auburn.lon <- All.Cities$long[All.Cities$city == "Auburn"]

# Generate Empty Dataframe
Cities.Distance.NewYork <- NULL # create a null object

# Run Function
Distance.NewYork <- distance(NewYork.lat,NewYork.lon,Auburn.lat,Auburn.lon)

# Combine Empty and Output Dataframe
Cities.Distance.NewYork <- rbind.data.frame(Cities.Distance.NewYork, Distance.NewYork)

# Edit Column Name
colnames(Cities.Distance.NewYork) <- "km_distance"

# Print Values
print(Cities.Distance.NewYork)
```

### Question 6 - For Loop
Now, use your function within a for loop to calculate the distance between all other Cities.nm in the data. 

```{r Question 6, eval=FALSE}
# Select All Cities That Are Not Auburn
Citites <- All.Cities[All.Cities$city != "Auburn",]

# Create a vector of City Names
Citites.names <- unique(Citites$city)

# Generate Empty Dataframe
Cities.Distance.All <- NULL # create a null object

# Create for Loop that runs through all Cities and Calculates Distance to Auburn
for (i in seq_along(Citites.names)) {
  Cities.Distance.Single <- distance(Citites$lat[i],
                                   Citites$long[i],
                                   Auburn.lat,
                                   Auburn.lon)
  Cities.Distance.All <- rbind.data.frame(Cities.Distance.All, Cities.Distance.Single)
}

# Create a Dataframe with Similar Observation Length as All.Citites
Citites.Two <- rep("Auburn",39)

# Combine Three Dataframes
Cities.Distance.All <- cbind(Citites.names,City.Two.Name,Cities.Distance.All)

# Edit Column Names
colnames(Cities.Distance.All) <- c("City1", "City2", "km_distance")

# Print Values
print(Cities.Distance.All)
```

