---
title: "Coding Challege 4"
author: "Sam Donohoo"
date: "2025-01-27"
output:
  md_document:
    variant: gfm
  html_document:
    toc_float: true
  word_document:
    toc: true
  pdf_document:
    toc: true
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = FALSE, message = FALSE)
```

```{r, include=TRUE}
library(knitr)
library(ggplot2)
library(markdown)
```

# Question 1

**Explain a YAML Header:** A header that lists the title, author, date, and output formats for the markdown.

**Explain Literate Programming:** Coding organized into small sections to be more easily read.

# Question 2

#### Take the code you wrote for coding challenge 3, question 5, and incorporate it into your R markdown file.

Data from Noel, Z.A., Roze, L.V., Breunig, M., Trail, F. 2022. Endophytic fungi as promising biocontrol agent to protect wheat from Fusarium graminearum head blight. Plant Disease. https://doi.org/10.1094/PDIS-06-21-1253-RE

### Individual Graphs
```{r Question 2, echo=TRUE}
# Load in Libraries
library(tidyverse)
library(ggpubr)
library(ggrepel)

# Make a color-blind friendly palette
cbbPalette <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7", "#000000")

# Load in Data
myco <- read.csv("MycotoxinData.csv", header = TRUE, na.strings = "na")

# Reorder by Treatment Level
myco$Treatment <- factor(myco$Treatment, levels = c("NTC", "Fg", "Fg + 37", "Fg + 40", "Fg + 70"))

#### Graph the effects of DON by Treatment
myco.don.Q1 <- ggplot(myco, aes(x = Treatment, y = DON, color = Cultivar, fill = Cultivar)) +
  geom_boxplot(outlier.shape = NA, color = "black") +
  geom_point(aes(fill=Cultivar), pch=21, color = "black", 
             position=position_jitterdodge(dodge.width=0.85), alpha = 0.6) + 
  scale_color_manual(values = cbbPalette) +
  scale_fill_manual(values = cbbPalette) + 
  ylab("DON (ppm)") +
  xlab("") +
  theme_classic() +
  facet_wrap(~Cultivar)

myco.don.Q1

### Graph the effects of X15ADON by Treatment
myco.don.Q3.1 <- ggplot(myco, aes(x = Treatment, y = X15ADON, color = Cultivar, fill = Cultivar)) +
  geom_boxplot(outlier.shape = NA, color = "black") +
  geom_point(aes(fill=Cultivar), pch=21, color = "black", 
             position=position_jitterdodge(dodge.width=0.85), alpha = 0.6) + 
  scale_color_manual(values = cbbPalette) +
  scale_fill_manual(values = cbbPalette) + 
  ylab("15ADON (ppm)") +
  xlab("") +
  theme_classic() +
  facet_wrap(~Cultivar)

myco.don.Q3.1

### Graph the effects of MassperSeed_mg by Treatment
myco.don.Q3.2 <- ggplot(myco, aes(x = Treatment, y = MassperSeed_mg, color = Cultivar, fill = Cultivar)) +
  geom_boxplot(outlier.shape = NA, color = "black") +
  geom_point(aes(fill=Cultivar), pch=21, color = "black", 
             position=position_jitterdodge(dodge.width=0.85), alpha = 0.6) + 
  scale_color_manual(values = cbbPalette) +
  scale_fill_manual(values = cbbPalette) + 
  ylab("Seed Mass (mg)") +
  xlab("") +
  theme_classic() +
  facet_wrap(~Cultivar)

myco.don.Q3.2
```

### T-Tests
```{r}
# Question 1 + Ttest
myco.don.Q1.Ttest <- myco.don.Q1 + 
  geom_pwc(aes(group = Treatment), method = "t_test", label = "{p.adj.signif}")

myco.don.Q1.Ttest

# Question 3 - 15ADON + Ttest
myco.don.Q3.1.Ttest <- myco.don.Q3.1 + 
  geom_pwc(aes(group = Treatment), method = "t_test", label = "{p.adj.signif}")

myco.don.Q3.1.Ttest

# Question 3 - MassperSeed_mg + Ttest

myco.don.Q3.2.Ttest <- myco.don.Q3.2 + 
  geom_pwc(aes(group = Treatment), method = "t_test", label = "{p.adj.signif}")

myco.don.Q3.2.Ttest
```

### Combined Figure
```{r}
myco.don.Ttest <- ggarrange(
  myco.don.Q1.Ttest,  # First plot: Q1 Ttest
  myco.don.Q3.1.Ttest,  # Second plot: Q3.1 Ttest
  myco.don.Q3.2.Ttest,  # Third plot: Q3.2 Ttest
  labels = "AUTO",  # Automatically label the plots (A, B, C, etc.)
  nrow = 1,  # Arrange the plots in 3 rows
  ncol = 3,  # Arrange the plots in 1 column
  legend = FALSE  # Do not include a legend in the combined figure
)

myco.don.Ttest
```

